---
title: "Impute missing value"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(VIM)
library(tidyverse)
library(stringr)
library(data.table)
```


Import data that includes all columns from 3 statements
```{r}
df <- read.csv('capstone_repository/data/df.csv')
allcol <- read.csv('capstone_repository/data/allColumns.csv')
```

Proportion of missing value in each column 
```{r}
prop_missing <- sort(round(colMeans(!is.na(df))*100,2), decreasing = T)
```

Operating Income
- 总共4个missing
```{r}
#check number of missing value
sum(is.na(df[,'Operating.Income..Loss.']))
#find 10 variables with the highest absolute correlation
cor <- abs(cor(subset(df[sapply(df, is.numeric)], select = -Operating.Income..Loss.), df$Operating.Income..Loss., use = 'complete.obs'))
high_cor <- head(cor[order(cor[,1],decreasing = T),],10)
high_cor
```


```{r}
#check the proporting of non-missing values in this column when the imputed col is missing
sort(round(colMeans(!is.na(df[is.na(df$Operating.Income..Loss.),]))*100,2), decreasing = T)[names(high_cor)]
#impute
df <- regressionImp(Operating.Income..Loss.~Pretax.Income..Loss...Adj.+Income..Loss..from.Continuing.Operations+Pretax.Income..Loss.+ Income.Tax..Expense..Benefit..Net + Net.Income..Common. + Dividends.Paid, data = df, imp_var= F)
```

Total Current Asset
- 4 missing 
```{r}
#check number of missing values
sum(is.na(df$Total.Current.Assets))
#find 10 variables with the highest absolute correlation
cor <- abs(cor(subset(df[sapply(df, is.numeric)], select = -Total.Current.Assets), df$Total.Current.Assets, use = 'complete.obs'))
high_cor <- head(cor[order(cor[,1],decreasing = T),],10)
high_cor
```


```{r}
#check the proporting of non-missing values in this column when the imputed col is missing
sort(round(colMeans(!is.na(df[is.na(df$Total.Current.Assets),]))*100,2), decreasing = T)[names(high_cor)]
#impute
df <- regressionImp(Total.Current.Assets~ Total.Assets + Total.Liabilities...Equity + Total.Current.Liabilities + Total.Liabilities, data = df, imp_var= F)
```

Total Equity
- 1 missing 
- TOtal Equity = Total Equity&Liability - Total Liabilties
- This company doesn't have equity, use 0 to fill
```{r}
sum(is.na(df$Total.Equity))
df[is.na(df$Total.Equity),]
df$Total.Equity <- replace_na(df$Total.Equity, replace = 0)
```

Depreciation & Amortization
- 233 missing values
- need to create a imp_var
```{r}
#check number of missing values
sum(is.na(df$Depreciation...Amortization))
#find 10 variables with the highest absolute correlation
cor <- abs(cor(subset(df[sapply(df, is.numeric)], select = -Depreciation...Amortization), df$Depreciation...Amortization, use = 'complete.obs'))
high_cor <- head(cor[order(cor[,1],decreasing = T),],10)
high_cor
```


```{r}
#check the proporting of non-missing values in this column when the imputed col is missing
sort(round(colMeans(!is.na(df[!is.na(df$Depreciation...Amortization),]))*100,2), decreasing = T)[names(high_cor)]
#impute
df <- regressionImp(Depreciation...Amortization~ Total.Assets +  Total.Liabilities...Equity + Total.Liabilities + Total.Current.Liabilities, data = df, imp_var= T)
```


Long Term Debt
- 2893 missing values
- Can't use Depreciation & Amortization because the variabe itself is imputed
```{r}
#check number of missing values
sum(is.na(df$Long.Term.Debt))
#find 10 variables with the highest absolute correlation
cor <- abs(cor(subset(df[sapply(df, is.numeric)], select = -Long.Term.Debt), df$Long.Term.Debt, use = 'complete.obs'))
high_cor <- head(cor[order(cor[,1],decreasing = T),],15)
high_cor
```


```{r}
#check the proporting of non-missing values in this column when the imputed col is missing
sort(round(colMeans(!is.na(df[is.na(df$Long.Term.Debt),]))*100,2), decreasing = T)[names(high_cor)]
#impute
df <- regressionImp(Long.Term.Debt~ Dividends.Paid + Total.Liabilities, data = df, imp_var= T)
```


Revenue
- 206 missing value
- total current asset is imputed, and thus not used
```{r}
#check number of missing values
sum(is.na(df$Revenue))
#find 10 variables with the highest absolute correlation
cor <- abs(cor(subset(df[sapply(df, is.numeric)], select = -Revenue), df$Revenue, use = 'complete.obs'))
high_cor <- head(cor[order(cor[,1],decreasing = T),],10)
high_cor
```


```{r}
#check the proporting of non-missing values in this column when the imputed col is missing
sort(round(colMeans(!is.na(df[is.na(df$Revenue),]))*100,2), decreasing = T)[names(high_cor)]
#impute
df <- regressionImp(Revenue~ Total.Current.Liabilities + Total.Liabilities + Total.Assets + Total.Liabilities...Equity, data = df, imp_var= T)

```

Income Tax Expense
- 737 missing values

```{r}
#check number of missing values
sum(is.na(df$Income.Tax..Expense..Benefit..Net))
#find 10 variables with the highest absolute correlation
cor <- abs(cor(subset(df[sapply(df, is.numeric)], select = -Income.Tax..Expense..Benefit..Net), df$Income.Tax..Expense..Benefit..Net, use = 'complete.obs'))
high_cor <- head(cor[order(cor[,1],decreasing = T),],10)
high_cor
```


```{r}
#check the proporting of non-missing values in this column when the imputed col is missing
sort(round(colMeans(!is.na(df[!is.na(df$Income.Tax..Expense..Benefit..Net),]))*100,2), decreasing = T)[names(high_cor)]
#impute
df <- regressionImp(Income.Tax..Expense..Benefit..Net~ Pretax.Income..Loss. + Pretax.Income..Loss...Adj. + Operating.Income..Loss. + Income..Loss..from.Continuing.Operations + Net.Income..Common. + Total.Assets, data = df, imp_var= T)
```


Shares Diluted
- 75 missing
```{r}
#check number of missing values
sum(is.na(df$Shares..Diluted.))
#find 10 variables with the highest absolute correlation
cor <- abs(cor(subset(df[sapply(df, is.numeric)], select = -Shares..Diluted.), df$Shares..Diluted., use = 'complete.obs'))
high_cor <- head(cor[order(cor[,1],decreasing = T),],10)
high_cor
```


```{r}
#check the proporting of non-missing values in this column when the imputed col is missing
sort(round(colMeans(!is.na(df[!is.na(df$Shares..Diluted.),]))*100,2), decreasing = T)[names(high_cor)]
#impute
df <- regressionImp(Shares..Diluted.~ Total.Assets + Total.Liabilities...Equity, data = df, imp_var= F)
```


Inventories
- 4174 missing
- About 35% of technology companies don't have inventories 
```{r}
#check number of missing values
sum(is.na(df$Inventories))
df1 <- as.data.table(subset(df, select = c('Inventories', 'Sector'，'Industry')))
df1[, .(prop_missing = mean(is.na(Inventories))),by = Sector][order(-rank(prop_missing))]


```


```{r}
#find 10 variables with the highest absolute correlation
cor <- abs(cor(subset(df[sapply(df, is.numeric)], select = -Inventories), df$Inventories, use = 'complete.obs'))
high_cor <- head(cor[order(cor[,1],decreasing = T),],10)
high_cor
```


```{r}
#check the proporting of non-missing values in this column when the imputed col is missing
sort(round(colMeans(!is.na(df[!is.na(df$Inventories),]))*100,2), decreasing = T)[names(high_cor)]
#impute
df <- regressionImp(Inventories~ Total.Assets + Total.Liabilities...Equity + Total.Current.Liabilities + Total.Liabilities , data = df, imp_var= T)
```


```{r}
write.csv(df, file = 'capstone_repository/data/df2.csv')
```

