---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Setup
```{r}
library(tidyverse)
library(DataExplorer)
library(ggplot2)
library(data.table)
```

Import data 
```{r}
df <- read.csv("capstone_repository/data/ratios_return.csv")
dim(df)#10788    35
colnames(df)
```

Keeps rows in Technology Sector
```{r}
tech <- subset(df, Sector == 'Technology')
dim(tech)# 2085   35
```






#####################################################################Operating.Profit.Margin############################################

Outliers on the left side 
```{r}
summary(tech$Operating.Profit.Margin)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#-12.07678   0.01599   0.08713   0.04513   0.17861   0.62948 

#check distribuion
ggplot(tech)+ geom_histogram(aes(x=Operating.Profit.Margin)) 
```


After removing observation smaller than -1, the distribution is slightly left skewed
most observations are around 0
we may want to try some transformations
```{r}
#arbitrily choose a subset 
sum(tech$Operating.Profit.Margin <  -1) #23

ggplot(tech[tech$Operating.Profit.Margin > -1,]) + geom_histogram(aes(x = Operating.Profit.Margin))
```



Cube transformation doesn't work
```{r}
#function for cube transformation
#Math.cbrt <- function(x) {
#    sign(x) * abs(x)^(1/3)
#}

#tech$Operating.Profit.Margin.cube <- Math.cbrt(tech$Operating.Profit.Margin)
#ggplot(tech[tech$Operating.Profit.Margin > -1,]) + geom_histogram(aes(x = Operating.Profit.Margin.cube)) #doesn't work
```



log transformation, doesn't help either 
```{r}
# #Math.log <- function(x){
#   min_x <- min(x)
#   log(x + abs(min_x) + 0.1)
# }
# 
# 
# tech$Operating.Profit.Margin.log <- Math.log(tech$Operating.Profit.Margin)
# summary(tech$Operating.Profit.Margin.log)
# ggplot(tech[tech$Operating.Profit.Margin > -1,]) + geom_histogram(aes(x = Operating.Profit.Margin.log)) #doesn't work
```


So we will not do any transfrmation on this data but instead remove several outliers. 
```{r}
sd <- sd(tech$Operating.Profit.Margin) #0.4783343
mean <- mean(tech$Operating.Profit.Margin) # 0.04513398
ci <- c(mean - 3*sd, mean + 3*sd) #-1.389869  1.480137
sum(!tech$Operating.Profit.Margin %between% ci)#15 outliers

#plot the histogram after removing outliers 
ggplot(tech[tech$Operating.Profit.Margin %between% ci,]) + geom_histogram(aes(x = Operating.Profit.Margin))
```


Ticker and Year for outliers 
```{r}
tech$Operating.Profit.Margin.out <- !tech$Operating.Profit.Margin %between% ci
sum(tech$Operating.Profit.Margin.out)
```
Summary:
1. No transformation
2. those outliers are outside 3 standard deviation from the mean 








####################################################################  Return.On.Capital#################################################

pretty right skewed 
```{r}
summary(tech$Return.On.Capital)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#-6.2459  0.1449  0.3547  0.6844  0.7674 40.5840 

#check distribuion
ggplot(tech)+ geom_histogram(aes(x=Return.On.Capital)) 
```





After removing observatins larger than 10, the distribution is still pretty right skewed 
```{r}
#even after removing outliers the data is still pretty right skewed 
ggplot(tech[tech$Return.On.Capital < 10,]) + geom_histogram(aes(x = Return.On.Capital))
```

log transformation doesn't help
```{r}
#tech$Return.On.Capital.log <- Math.log(tech$Return.On.Capital)
#ggplot(tech[tech$Return.On.Capital < 10,]) + geom_histogram(aes(x = Return.On.Capital.log))
```




still has 35 outliers after transformation in terms of standard deviation
The correlaition is improved but still minor
but the distribution looks pretty good, so we are not removing any outliers
```{r}
tech$Return.On.Capital.cube <- Math.cbrt(tech$Return.On.Capital)

sd_cube <- sd(tech$Return.On.Capital.cube)
mean <- mean(tech$Return.On.Capital.cube)
ci <- c(mean - 3*sd_cube, mean + 3* sd_cube)
sum(!tech$Return.On.Capital.cube %between% ci) #still has 36 outliers 

#plot distribution
ggplot(tech[tech$Return.On.Capital.cube  %between% ci,]) + geom_histogram(aes(x = Return.On.Capital.cube))

```

Identify outliers 
```{r}
tech$Return.On.Capital.out <- !tech$Return.On.Capital.cube  %between% ci
sum(tech$Return.On.Capital.out)
```
Summary 
1. cube transformation
2. removing outliers outside3 sd






#######################################Return on Equity########################################

```{r}
summary(tech$Return.On.Equity)
#    Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
#-10.9390   0.3372   0.6366   1.7279   1.2912 964.7980 

#check distribuion
ggplot(tech)+ geom_histogram(aes(x=Return.On.Equity)) 
```




after removing several observations still 0
```{r}
sum(tech$Return.On.Equity > 10) #37
ggplot(tech[tech$Return.On.Equity < 10,])+ geom_histogram(aes(x=Return.On.Equity)) 
```




try log transformation, doesn't work well
```{r}
#tech$Return.On.Equity.log <-  Math.log(tech$Return.On.Equity)
#ggplot(tech[tech$Return.On.Equity < 10,])+ geom_histogram(aes(x=Return.On.Equity.log)) 
```



try cube transformation, works good
```{r}
tech$Return.On.Equity.cube <-  Math.cbrt(tech$Return.On.Equity)

ggplot(tech[tech$Return.On.Equity < 10,])+ geom_histogram(aes(x=Return.On.Equity.cube)) 
```


identify outliers
```{r}
sd <- sd(tech$Return.On.Equity.cube) # 0.5657205
mean <- mean(tech$Return.On.Equity.cube) #0.8724592
ci <- c(mean - 3*sd, mean + 3*sd) #-0.8247022  2.5696207
sum(!tech$Return.On.Equity.cube %between% ci)#30 outliers


ggplot(tech[tech$Return.On.Equity.cube %between% ci,]) + geom_histogram(aes(x = Return.On.Equity.cube))
```

```{r}
tech$Return.On.Equity.out <- !tech$Return.On.Equity.cube %between% ci
sum(tech$Return.On.Equity.out)
```




#############################################Receivable Turnover##########################################



```{r}
summary(tech$Receivable.Turnover)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#   0.2055    5.4420    6.9895   12.5850    9.3841 2829.0083 

#check distribuion
ggplot(tech)+ geom_histogram(aes(x=Receivable.Turnover)）
```


After removing outliers, still right skewed 
```{r}
sum(tech$Receivable.Turnover > 100) #16
ggplot(tech[tech$Receivable.Turnover < 100,])+ geom_histogram(aes(x=Receivable.Turnover))
```

Try log transformation, works pretty well
```{r}
tech$Receivable.Turnover.log <- Math.log(tech$Receivable.Turnover)
ggplot(tech[tech$Receivable.Turnover < 100,])+ geom_histogram(aes(x=Receivable.Turnover.log))
```

Compare with cube transformation, log works better 
```{r}
#tech$Receivable.Turnover.cube <- Math.cbrt(tech$Receivable.Turnover)
#ggplot(tech[tech$Receivable.Turnover < 100,])+ geom_histogram(aes(x=Receivable.Turnover.cube))
```



identify outliers
```{r}
sd <- sd(tech$Receivable.Turnover.log) #  0.6323837
mean <- mean(tech$Receivable.Turnover.log) #2.075199
ci <- c(mean - 3*sd, mean + 3*sd) #0.1780484 3.9723505
sum(!tech$Receivable.Turnover.log %between% ci)#47 outliers

ggplot(tech[tech$Receivable.Turnover.log %between% ci,])+ geom_histogram(aes(x=Receivable.Turnover.log))
```

```{r}
tech$Receivable.Turnover.out <- !tech$Receivable.Turnover.log %between% ci
sum(tech$Receivable.Turnover.out) #47
```




#################################################### Book.to.Market Ratio ############################################################

right skewed
```{r}
summary(tech$Book.to.Market.Ratio)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#   0.0042    0.1965    0.3437    2.8135    0.6151 1326.5407 

#check distribuion
ggplot(tech)+ geom_histogram(aes(x=Book.to.Market.Ratio)) 
```


removing several outliers, still right skewed distribution
```{r}
#the first histogram shows to be pretty rights skewed, so we set a arbitrary cut off at 10
sum(tech$Book.to.Market.Ratio > 10) #11

#even after removing observations > 10 the data is still pretty right skewed 
ggplot(tech[tech$Book.to.Market.Ratio <10,]) + geom_histogram(aes(x = Book.to.Market.Ratio))
```


Try cube transformation, doesn't work so well
```{r}
#tech$Book.to.Market.Ratio.cube <- Math.cbrt(tech$Book.to.Market.Ratio)
#ggplot(tech) + geom_histogram(aes(x = Book.to.Market.Ratio.cube))
```



Try log transformation, works better, with wider range
after removing observations 3sd more from the mean, the distribution is very symmetric 
```{r}
tech$Book.to.Market.Ratio.log <- Math.log(tech$Book.to.Market.Ratio)
ggplot(tech) + geom_histogram(aes(x = Book.to.Market.Ratio.log))


sd <- sd(tech$Book.to.Market.Ratio.log) # 0.7624426
mean <- mean(tech$Book.to.Market.Ratio.log) #-0.736918
ci <- c(mean - 3*sd, mean + 3*sd) #-3.024246  1.550410
sum(!tech$Book.to.Market.Ratio.log %between% ci)#14 outliers


ggplot(tech[tech$Book.to.Market.Ratio.log %between% ci,]) + geom_histogram(aes(x = Book.to.Market.Ratio.log))
```

```{r}
tech$Book.to.Market.Ratio.out <- !tech$Book.to.Market.Ratio.log %between% ci
sum(tech$Book.to.Market.Ratio.out)
```
Summary
1. log transformation 
2. remove 14 outliers 3 sd away from mean

####################################################################write files #############################



```{r}
#join with stock return, set up
stock_return <- read.csv("capstone_repository/data/stock_return.csv")[-1]

ratios_t <- tech[c("Ticker","SimFinId", "Company.Name", "Sector", "Industry", "Fiscal.Year", "Operating.Profit.Margin","Operating.Profit.Margin.out","Return.On.Capital", "Return.On.Capital.cube", "Return.On.Capital.out", "Return.On.Equity", "Return.On.Equity.cube", "Return.On.Equity.out", "Book.to.Market.Ratio", "Book.to.Market.Ratio.log", "Book.to.Market.Ratio.out" ,"Receivable.Turnover", "Receivable.Turnover.log","Receivable.Turnover.out")]


setnames(ratios_t, old = c("Return.On.Capital.cube", "Return.On.Equity.cube", "Book.to.Market.Ratio.log", "Receivable.Turnover.log"), new = c("Return.On.Capital.t", "Return.On.Equity.t", "Book.to.Market.Ratio.t", "Receivable.Turnover.t"))

stock_return$Year.adj <- stock_return$Year - 1L
stock_return$Ticker <- as.character(stock_return$Ticker)
ratios_t$Ticker <- as.character(ratios_t$Ticker)
yimeng_transformed <- left_join(ratios_t, stock_return, by = c("Fiscal.Year" = "Year", "Ticker" = 'Ticker', "SimFinId" = 'SimFinId'))

```




```{r}
#check result
cor(yimeng_transformed$Operating.Profit.Margin, yimeng_transformed$Return3)
cor(yimeng_transformed$Return.On.Capital.t, yimeng_transformed$Return3)
cor(yimeng_transformed$Return.On.Equity.t, yimeng_transformed$Return3)
cor(yimeng_transformed$Receivable.Turnover.t, yimeng_transformed$Return3)
cor(yimeng_transformed$Book.to.Market.Ratio.t, yimeng_transformed$Return3)

# [1] 0.1294372
# [1] 0.07710282
# [1] 0.07230884
# [1] 0.09234118
# [1] -0.2368317
```

############################################################write files to local####################################
```{r}

```





