---
title: "Calculate ratios"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Import libraries
```{r, warining = F}
library(tidyverse)
library(ggplot2)
library(VIM)
library(data.table)
```


import statement data
```{r}
df <- fread('capstone_repository/data/allcol.csv', drop = 1)
dim(df)

#no missing value in the original statement data 
colMeans(!is.na(df))
```

Set up change id variables to character type, identify numeric columsn and categorical columns and logical columns
```{r}
names(df)
df$SimFinId <- as.character(df$SimFinId)
df$IndustryId <- as.character(df$IndustryId)

#create a copy for ratio calculation
ratio <- copy(df)
```

Current Ratio: Total Current Asset/ Total Current Liabilities 
```{r}
#check if there are zeros or negative value that interrupt the calculation
#no negative values
summary(ratio$Total.Current.Assets)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#0.000e+00 1.646e+08 6.461e+08 3.452e+09 2.324e+09 2.031e+11
sum(ratio$Total.Current.Assets == 0) #2

summary(ratio$Total.Current.Liabilities)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#0.000e+00 1.646e+08 6.461e+08 3.452e+09 2.324e+09 2.031e+11
sum(ratio$Total.Current.Liabilities == 0) #2


#start calculation
ratio <- ratio %>% mutate(Current.Ratio = ifelse(Total.Current.Liabilities==0, NA, round(Total.Current.Assets/Total.Current.Liabilities,6)))


#check result
#2 invalid(might need to drop later)
#1 zero
head(ratio[,c('Total.Current.Assets', 'Total.Current.Liabilities','Current.Ratio')])
sum(is.na(ratio$Current.Ratio))
sum(ratio$Current.Ratio == 0, na.rm = T)
```

Quick Ratio: (Cash equivalent + Account receivable)/Total Current Liabilities
```{r}
#check if there are zeros or negative value that interrupt the calculation
summary(ratio$Cash..Cash.Equivalents...Short.Term.Investments)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#0.000e+00 3.780e+07 1.638e+08 1.231e+09 6.376e+08 1.338e+11 
summary(ratio$Accounts...Notes.Receivable)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#0.000e+00 3.748e+07 1.793e+08 8.664e+08 6.740e+08 6.555e+10 
summary(ratio$Total.Current.Liabilities)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#0.000e+00 6.887e+07 3.601e+08 2.576e+09 1.635e+09 2.308e+11 

#start calculation
ratio <- ratio %>% mutate(Quick.Ratio = ifelse(Total.Current.Liabilities==0, NA, round((Cash..Cash.Equivalents...Short.Term.Investments + Accounts...Notes.Receivable)/Total.Current.Liabilities,6)))

#check result
#2 invalid
#0 zero
head(ratio[,c('Cash..Cash.Equivalents...Short.Term.Investments', 'Accounts...Notes.Receivable','Total.Current.Liabilities','Quick.Ratio')])
sum(is.na(ratio$Quick.Ratio))
mean(ratio$Quick.Ratio == 0, na.rm=T)
```

Net working capital to asset ratio: (Total Current Asset - Total Current Liabilities)/ Total Current Asset
```{r}
#start calculation
ratio <- ratio %>% mutate(NWC.to.Asset.Ratio = ifelse(Total.Current.Assets==0, NA, round((Total.Current.Assets-Total.Current.Liabilities)/Total.Current.Assets,6)))


#check result
#2 invalid
#2380 zero(can be negative)
head(ratio[,c('Total.Current.Assets', 'Total.Current.Liabilities','NWC.to.Asset.Ratio')])
sum(is.na(ratio$NWC.to.Asset.Ratio))
sum(ratio$NWC.to.Asset.Ratio == 0, na.rm = T)
sum(ratio$NWC.to.Asset.Ratio < 0, na.rm = T)
```

Cash Ratio: Cash Equivalent/ Total Current Liabilities
```{r}
#start calculation
ratio <- ratio %>% mutate(Cash.Ratio = ifelse(Total.Current.Liabilities==0, NA, round(Cash..Cash.Equivalents...Short.Term.Investments/Total.Current.Liabilities,6)))


#check result
#2 invalid
#57 zero
head(ratio[,c('Cash..Cash.Equivalents...Short.Term.Investments', 'Total.Current.Liabilities','Cash.Ratio')])
sum(is.na(ratio$Cash.Ratio))
sum(ratio$Cash.Ratio == 0, na.rm = T)
```


Long term Debt Ratio: Long term debt/(Long Term Debt + Total Equity)
```{r}
#check if there are zeros or negative value that interrupt the calculation
summary(ratio$Long.Term.Debt)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-3.480e+05  2.263e+08  5.630e+08  2.870e+09  2.207e+09  1.662e+11 
#check if the negative values are imputed
sum(ratio$Long.Term.Debt <0 ) #2 negative,
#Most negative liabilities are created in error, so their presence indicates problems with the underlying accounting system. need to make it NA

summary(ratio$Total.Equity)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.453e+10  1.356e+08  7.464e+08  4.045e+09  2.894e+09  3.525e+11
sum(ratio$Total.Equity < 0, na.rm = T) #717
#Owner's equity can be reported as a negative on a balance sheet; however, if the owner's equity is negative, the company owes more than it is worth at that point in time.
head(ratio[ratio$Total.Equity < 0, c('Total.Assets', 'Total.Liabilities', 'Total.Equity')])
#the original data is correct, we don't need to process the data 


#start calculation
ratio <- ratio %>% mutate(Long.Term.Debt.Ratio = ifelse(((Long.Term.Debt + Total.Equity)==0) | Long.Term.Debt <0 , NA, round(Long.Term.Debt/(Long.Term.Debt + Total.Equity),6)))

#check result
#2 NA
#275 zero
#101 negative
head(ratio[,c('Total.Equity', 'Long.Term.Debt','Long.Term.Debt.Ratio')])
sum(is.na(ratio$Long.Term.Debt.Ratio))
sum(ratio$Long.Term.Debt.Ratio == 0, na.rm = T)
sum(ratio$Long.Term.Debt.Ratio < 0, na.rm = T)
```

Liabilities to Equity Ratio: Total Liabilities/ Total Equity
```{r}
#check if there are zeros or negative value that interrupt the calculation
summary(ratio$Total.Liabilities)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.743e+09  1.363e+08  1.096e+09  6.952e+09  5.029e+09  3.553e+11 
sum(ratio$Total.Liabilities <0, na.rm = T) # 15
#Most negative liabilities are created in error, so their presence indicates problems with the underlying accounting system. We want to make it NA

summary(ratio$Total.Equity)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.453e+10  1.356e+08  7.464e+08  4.045e+09  2.894e+09  3.525e+11 
sum(ratio$Total.Equity <0 ) #717, negative equity is red flag of the over-leveraging, keep it

#start calculation
ratio <- ratio %>% mutate(Liabilities.to.Equity.Ratio = ifelse(Total.Equity==0 | Total.Liabilities < 0, NA, round(Total.Liabilities/Total.Equity,6)))



#check result
#26 NA
#2 zero
#717 negative 
head(ratio[,c('Total.Equity', 'Total.Liabilities','Liabilities.to.Equity.Ratio')])
sum(is.na(ratio$Liabilities.to.Equity.Ratio))
sum(ratio$Liabilities.to.Equity.Ratio == 0, na.rm = T)
sum(ratio$Liabilities.to.Equity.Ratio < 0, na.rm = T)
```


Total Debt Ratio: Total Liabilities/Total Assets
```{r}
#check if there are zeros or negative value that interrupt the calculation
summary(ratio$Total.Liabilities)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.743e+09  1.363e+08  1.096e+09  6.952e+09  5.029e+09  3.553e+11 
sum(ratio$Total.Liabilities<0) #15
#Assets all positive
summary(ratio$Total.Assets)


#start calculation
ratio <- ratio %>% mutate(Total.Debt.Ratio = ifelse(Total.Assets==0| Total.Liabilities <0, NA, round(Total.Liabilities/Total.Assets,6)))


#check result
#15 NA
#2 zero
#0 negative (mostly due to negative Total.Equity )
head(ratio[,c('Total.Assets', 'Total.Liabilities','Total.Debt.Ratio')])
sum(is.na(ratio$Total.Debt.Ratio))
sum(ratio$Total.Debt.Ratio == 0, na.rm = T)
sum(ratio$Total.Debt.Ratio < 0, na.rm = T)
```

Debt to Asset Ratio:(Long term Debt + Short Term Debt)/ Total Assets
```{r}
#check if there are zeros or negative value that interrupt the calculation

summary(ratio$Long.Term.Debt)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-3.480e+05  2.263e+08  5.630e+08  2.870e+09  2.207e+09  1.662e+11
sum(ratio$Long.Term.Debt <0) #2, need to make it NA
summary(ratio$Short.Term.Debt)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#0.000e+00 0.000e+00 3.536e+06 3.590e+08 8.896e+07 5.349e+10
#Assets all positive
summary(ratio$Total.Assets)


#start calculation
ratio <- ratio %>% mutate(Debt.to.Asset.Ratio = ifelse((Total.Assets==0| Long.Term.Debt <0), NA, round((Long.Term.Debt + Short.Term.Debt)/Total.Assets,6)))




#check result
#2 NA
#201 zero
#0
head(ratio[,c('Long.Term.Debt','Short.Term.Debt' ,'Total.Liabilities','Debt.to.Asset.Ratio')])
sum(is.na(ratio$Debt.to.Asset.Ratio))
sum(ratio$Debt.to.Asset.Ratio == 0, na.rm = T)
sum(ratio$Debt.to.Asset.Ratio < 0, na.rm = T)
```


Interest Coverage Ratio: Operating Income/Interest expense
```{r}
#check if there are zeros or negative value that interrupt the calculation
summary(ratio$Operating.Income..Loss.)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.347e+10  9.540e+06  1.618e+08  9.877e+08  6.887e+08  1.425e+11 
sum(ratio$Operating.Income..Loss. <0, na.rm = T)
#2591, it can be negative and very few of them are imputed 

summary(ratio$Interest.Expense..Net)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-9.737e+09 -1.088e+08 -3.017e+07 -1.311e+08 -1.989e+06  1.537e+09 
sum(ratio$Interest.Expense..Net < 0, na.rm = T)
#12407, expense is recorded as negative
#If you earned more income from interest that resulted in a positive interest gain
#if we have interest gain,we need to set Interest Coverage to a high value so it can pass a high interest coverage screen. We may want to use mean to fill these 

#总共分三种情况
#1. positive income & interst expense（正常）
#2. positive income & interest gain （用mean fill）
#3. negative income & interest expense (正常)
#4. negative income & interest earn （不好， 异常 drop）

nrow(ratio[ratio$Sector == 'Technology'& ratio$Interest.Expense..Net > 0 & ratio$Operating.Income..Loss. <0, ]) #186



#start calculation
ratio <- ratio %>% mutate(Interest.Coverage.Ratio = ifelse(Interest.Expense..Net>=0, NA, round(Operating.Income..Loss./(-1*Interest.Expense..Net),6)))
#assign mean value to those with interest gain
mean <- mean(ratio$Interest.Coverage.Ratio, na.rm=T) #50.039
ratio$Interest.Coverage.Ratio[ratio$Interest.Expense..Net>0 & ratio$Operating.Income..Loss. > 0] <- mean



#check result
#607 NA
#0 zero
#2025 negative (due to negative operating income)
head(ratio[,c('Operating.Income..Loss.', 'Interest.Expense..Net','Interest.Coverage.Ratio')])
sum(is.na(ratio$Interest.Coverage.Ratio))
sum(ratio$Interest.Coverage.Ratio == 0, na.rm = T)
sum(ratio$Interest.Coverage.Ratio < 0, na.rm = T)
```

Cash Coverage Ratio:(Operating Income + Depreciation)/Interest expense
```{r}
#check if there are zeros or negative value that interrupt the calculation
summary(ratio$Operating.Income..Loss.)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.347e+10  9.540e+06  1.618e+08  9.877e+08  6.887e+08  1.425e+11 
sum(ratio$Operating.Income..Loss. <0, na.rm = T)#2591

summary(ratio$Depreciation...Amortization)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.069e+09  1.158e+07  6.977e+07  4.150e+08  2.704e+08  3.876e+10 
sum(ratio$Depreciation...Amortization < 0, na.rm = T)#11
#Negative depreciation, on the other hand, accounts for the opposite process of an asset gaining value over time. but that's not cash flow, so we make it 0 here 
ratio$Depreciation...Amortization[ratio$Depreciation...Amortization < 0] <- 0

  

#三种情况
#1. 分子大于 0 & interst expense（正常）
#2. 分子大于 0 & interest gain （用mean fill）
#3. 分子小于 0 & interest expense (正常)
#4. 分子小于 0 & interest gain （不好， 异常 drop）

a <- ratio[ratio$Interest.Expense..Net > 0 & (ratio$Depreciation...Amortization + ratio$Operating.Income..Loss.) < 0 & ratio$Sector == 'Technology', ] # 155
b <- ratio[ratio$Sector == 'Technology'& ratio$Interest.Expense..Net > 0 & ratio$Operating.Income..Loss. <0, ]#86
result <- inner_join(a, b, on = c('Ticker' = 'Ticker','Fiscal.Year' = 'Fiscal.Year'))  #nrow  = 155 
#重合度高，可以drop

#start calculation
ratio <- ratio %>% mutate(Cash.Coverage.Ratio = ifelse(Interest.Expense..Net >= 0, NA, round((Operating.Income..Loss. + Depreciation...Amortization)/(-1*Interest.Expense..Net),6)))
mean <- mean(ratio$Cash.Coverage.Ratio, na.rm=T) #74.08185
ratio$Cash.Coverage.Ratio[ratio$Interest.Expense..Net>0 & (ratio$Depreciation...Amortization + ratio$Operating.Income..Loss.) > 0] <- mean


#check result
#544 NA
#0 zero
#1467 negative 
head(ratio[,c('Operating.Income..Loss.', 'Interest.Expense..Net','Depreciation...Amortization','Cash.Coverage.Ratio')])
sum(is.na(ratio$Cash.Coverage.Ratio))
sum(ratio$Cash.Coverage.Ratio == 0, na.rm = T)
sum(ratio$Cash.Coverage.Ratio < 0, na.rm = T)
```


After Tax Operating Income(Not a ratio)!!!: Revenue-Depreciaton-Income Tax expense
```{r}
summary(ratio$Revenue)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#0.000e+00 3.268e+08 1.527e+09 7.751e+09 5.466e+09 5.144e+11 
summary(ratio$Operating.Expenses)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.071e+11 -1.316e+09 -3.988e+08 -1.748e+09 -1.013e+08  1.110e+09 
sum(ratio$Operating.Expenses > 0 , na.rm = T)#4, should be problem, make it NA

summary(ratio$Depreciation...Amortization)      
#     Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.069e+09  1.158e+07  6.977e+07  4.150e+08  2.704e+08  3.876e+10 
sum(ratio$Depreciation...Amortization < 0, na.rm = T) #0

#start calculation
ratio <- ratio %>% mutate(After.Tax.Operating.Income = ifelse(Operating.Expenses >0,NA,Revenue - Depreciation...Amortization + Operating.Expenses))
head(ratio[,c("Revenue", "Depreciation...Amortization", "Operating.Expenses", "After.Tax.Operating.Income")])
sum(is.na(ratio$After.Tax.Operating.Income)) #4
sum(ratio$After.Tax.Operating.Income <0, na.rm=T) #1229
sum(ratio$After.Tax.Operating.Income ==0, na.rm =T) #1
```


Return On Assets: After Tax Operating Income/Total Assets
```{r}
summary(ratio$After.Tax.Operating.Income)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max.       NA's 
#-1.112e+10  1.174e+08  8.770e+08  5.589e+09  3.380e+09  4.188e+11          4 

summary(ratio$Total.Assets)      
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#2.360e+02 3.702e+08 2.062e+09 1.100e+10 8.207e+09 7.078e+11 


#start calculation
ratio <- ratio %>% mutate(Return.On.Assets = After.Tax.Operating.Income/Total.Assets)
sum(is.na(ratio$Return.On.Assets))#4
```



Return On Capital: After Tax Operating Income/(Long Term Debt + Total Equity)
```{r}
summary(ratio$After.Tax.Operating.Income)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max.       NA's 
#-1.112e+10  1.174e+08  8.770e+08  5.589e+09  3.380e+09  4.188e+11          4 

summary(ratio$Long.Term.Debt)      
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-3.480e+05  2.263e+08  5.630e+08  2.870e+09  2.207e+09  1.662e+11 
sum(ratio$Long.Term.Debt < 0) #2, make it NA

summary(ratio$Total.Equity)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.453e+10  1.356e+08  7.464e+08  4.045e+09  2.894e+09  3.525e+11 
sum(ratio$Total.Equity <0 , na.rm = T)#717

sum((ratio$Long.Term.Debt + ratio$Total.Equity) < 0 & ratio$Long.Term.Debt >= 0 & ratio$After.Tax.Operating.Income > 0 & ratio$Sector == 'Technology') 
#4钟情况
#1.分子正 & 分母正 （正常）
#2.分子正 & 分母负 （drop）21个
#3.分子负 & 分母正 （正常）
#4.分子负 & 分母负 （drop）5个


#start calculation
ratio <- ratio %>% mutate(Return.On.Capital = ifelse(((Long.Term.Debt + Total.Equity) <= 0) | Long.Term.Debt <0,NA,After.Tax.Operating.Income/(Long.Term.Debt + Total.Equity)))
sum(is.na(ratio$Return.On.Capital)) #118
sum(ratio$Return.On.Capital <0, na.rm=T) #1186
```

Return On Equity: After Tax Operating Income/Total Equity
```{r}
summary(ratio$After.Tax.Operating.Income)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.112e+10  1.175e+08  8.770e+08  5.589e+09  3.380e+09  4.188e+11 
sum(ratio$After.Tax.Operating.Income <0 , na.rm=T) #1229

summary(ratio$Total.Equity)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.453e+10  1.356e+08  7.464e+08  4.045e+09  2.894e+09  3.525e+11 
sum(ratio$Total.Equity < 0 , na.rm = T)
#717

#having positive profits and negative stockholder equity is more indicative of a potential future rebound than having negative profits and positive stockholder equity.So we need to differentiate the two 


sum(ratio$Total.Equity < 0 & ratio$After.Tax.Operating.Income < 0 & ratio$Sector == 'Technology') 
#4钟情况
#1.分子正 & 分母正 （正常）
#2.分子正 & 分母负 （drop）86个
#3.分子负 & 分母正 （正常）
#4.分子负 & 分母负 （drop）24个

#start calculation
#we need to create a indicator telling whether 
ratio <- ratio %>% mutate(Return.On.Equity = ifelse( Total.Equity <= 0,NA,After.Tax.Operating.Income/Total.Equity))


sum(is.na(ratio$Return.On.Equity)) #732
sum(ratio$Return.On.Equity <0, na.rm=T) #1080
```

Operating Profit Margin: Operating Income/ Revenue
```{r}
summary(ratio$Operating.Income..Loss.)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.347e+10  9.540e+06  1.618e+08  9.877e+08  6.887e+08  1.425e+11 
sum(ratio$Operating.Income..Loss. <0, na.rm = T) #2591

summary(ratio$Revenue)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#0.000e+00 3.268e+08 1.527e+09 7.751e+09 5.466e+09 5.144e+11 
sum(ratio$Revenue == 0 , na.rm = T) #53


#start calculation
ratio <- ratio %>% mutate(Operating.Profit.Margin = ifelse( Revenue == 0,NA,Operating.Income..Loss./Revenue))
sum(is.na(ratio$Operating.Profit.Margin)) #53
sum(ratio$Operating.Profit.Margin <0, na.rm=T) #2539
```


Net Profit Margin: Net Income/Revenue

```{r}
summary(ratio$Net.Income..Common.)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-2.280e+10 -6.136e+05  6.924e+07  5.606e+08  3.820e+08  5.953e+10 
sum(ratio$Net.Income..Common. <0, na.rm = T) #3608

summary(ratio$Revenue)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#0.000e+00 3.268e+08 1.527e+09 7.751e+09 5.466e+09 5.144e+11 
sum(ratio$Revenue == 0 , na.rm = T) #53


#start calculation
ratio <- ratio %>% mutate(Net.Profit.Margin = ifelse( Revenue == 0,NA,Net.Income..Common./Revenue))
sum(is.na(ratio$Net.Profit.Margin)) #53
sum(ratio$Net.Profit.Margin <0, na.rm=T) #3563
```

Asset Turnover; Revenue/Total Asset(year-start)
```{r, message warning}
#create a variable of year-start total asset
ratio <- as.data.table(ratio)
ratio[, Total.Assets.year.start := shift(Total.Assets, 1, 'lag'), by = Ticker]
head(ratio[,c("Ticker","Total.Assets", "Total.Assets.year.start")], 20)
sum(is.na(ratio$Total.Assets.year.start)) # 1750 missing
sum(ratio$Total.Assets.year.start ==0, na.rm = T) #0 zero
#complete

#calculate ratio
ratio <- ratio %>% mutate(Asset.Turnover = ifelse( Total.Assets.year.start == 0,NA,Revenue/Total.Assets.year.start))
sum(is.na(ratio$Asset.Turnover)) #1750
sum(ratio$Asset.Turnover <0, na.rm=T) #0
```



Receivable Turnover: Revenue/Account receivable year start
```{r, message warning}
#create a variable of year-start account receivable
ratio <- as.data.table(ratio)
ratio[, Accounts...Notes.Receivable.year.start := shift(Accounts...Notes.Receivable, 1, 'lag'), by = Ticker]

head(ratio[,c("Ticker","Accounts...Notes.Receivable", "Accounts...Notes.Receivable.year.start")], 20)
sum(is.na(ratio$Accounts...Notes.Receivable.year.start)) # 1750 missing
sum(ratio$Accounts...Notes.Receivable.year.start ==0, na.rm = T) #14 zero


#calculate ratio
ratio <- ratio %>% mutate(Receivable.Turnover = ifelse(Accounts...Notes.Receivable.year.start == 0, NA, Revenue/Accounts...Notes.Receivable.year.start))

sum(is.na(ratio$Receivable.Turnover)) #1764
sum(ratio$Receivable.Turnover <0, na.rm=T) #0
```


Inventory Turnover: Cost of Revenue/Inventory(year-start)
- we drop this because 
1. inventory has too many missing values 
2. Cost.of.Revenue has many postive values which is weird. 
```{r}
#create a variable of year-start inventory
ratio <- as.data.table(ratio)
ratio[, Inventories.year.start := shift(Inventories, 1, 'lag'), by = Ticker]

head(ratio[,c("Ticker","Inventories", "Inventories.year.start")], 20)
sum(is.na(ratio$Inventories.year.start)) # 1750 missing
sum(ratio$Inventories.year.start ==0, na.rm = T) #3634 zero

summary(ratio$Cost.of.Revenue)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-3.853e+11 -3.051e+09 -7.931e+08 -5.075e+09 -1.251e+08  3.946e+09 
sum(ratio$Cost.of.Revenue >0 , na.rm = T) #626
```


Book to Market Ratio: Stcok Price(year-end) * Shares(basic)/ Total Equity 
```{r}
#import stock price data
stock <- fread('simfin_data/us-shareprices-daily.csv')
stock$Year <- str_split_fixed(stock$Date,"-", n=3)[,1]
stock$Date2 <- str_remove_all(stock$Date, "-")
head(stock)

#find the year-end stock price 
end_price <- stock %>% group_by(Ticker, Year) %>% mutate(year_end = max(Date2)) %>% filter(Date2 == year_end)%>% select(Ticker, SimFinId, Date,Date2, Year, Close)
head(end_price,20) 
dim(end_price)

#join it with statement data 
ratio$Fiscal.Year <- as.character(ratio$Fiscal.Year)
end_price$SimFinId <- as.character(end_price$SimFinId)
ratio <- left_join(ratio,end_price, by= c("Ticker", "SimFinId", "Fiscal.Year" = "Year"))
sum(is.na(ratio$Close)) #777 missing, all because these comapnies are missing in stock price dataset

head(allcol)
#explore
summary(ratio$Close)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#      0      14      31    2603      58 6818748     772 
summary(ratio$Shares..Diluted.)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#1.400e+01 2.822e+07 7.300e+07 2.478e+08 2.119e+08 8.996e+09
summary(ratio$Total.Equity)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.453e+10  1.356e+08  7.464e+08  4.045e+09  2.894e+09  3.525e+11 
sum(ratio$Total.Equity==0, na.rm=T)#11
sum(ratio$Total.Equity<0) #717


summary(ratio$Close)
summary(ratio$Total.Equity)
summary(ratio$Shares..Diluted.)

#start calculation 
ratio <- ratio %>% mutate(Book.to.Market.Ratio = ifelse(Close ==0 , NA, (Total.Equity/Shares..Diluted.)/Close))

sum(is.na(ratio$Book.to.Market.Ratio)) #777
sum(ratio$Book.to.Market.Ratio <0, na.rm=T) #573
```

Earning Per Share: Net Income/ Shares
```{r}
summary(ratio$Net.Income..Common.)#some are negative
summary(ratio$Shares..Diluted.)

#start calculation 
ratio <- ratio %>% mutate(Earning.Per.Share = ifelse(Shares..Diluted. ==0 , NA, Net.Income..Common./Shares..Diluted.))

sum(is.na(ratio$Earning.Per.Share)) #0
sum(ratio$Earning.Per.Share <0, na.rm=T) #3608
```


Sales Per Share: Revenue/ Shares
```{r}
summary(ratio$Revenue)
summary(ratio$Shares..Diluted.)

#start calculation 
ratio <- ratio %>% mutate(Sales.Per.Share = ifelse(Shares..Diluted. ==0 , NA, Revenue/Shares..Diluted.))

sum(is.na(ratio$Sales.Per.Share)) #0
sum(ratio$Sales.Per.Share <0, na.rm=T) #0
```


Pricing to Earning Ratio: Stock Price/ Earning Per Share
```{r}
summary(ratio$Close)
summary(ratio$Earning.Per.Share)

#start calculation 
ratio <- ratio %>% mutate(Price.to.Earning.Ratio = ifelse(Earning.Per.Share ==0 , NA, Close/Earning.Per.Share))

sum(is.na(ratio$Price.to.Earning.Ratio)) #778
sum(ratio$Price.to.Earning.Ratio <0, na.rm=T) #3189
```


select only ratio columns, and drop rows with any missing value
```{r}
names(ratio)
final <- ratio[,c(1:8, 38:47,49:53, 55, 57, 61:65)]
final_withoutmissing <- na.omit(final)
dim(final)#[1] 13992    30
dim(final_withoutmissing) #10810 31
#we have 2297 rows with no missing values in technology sector 
nrow(final_withoutmissing[final_withoutmissing$Sector == 'Technology',]) # 2091

```

export data
```{r}
fwrite(final_withoutmissing, file = "capstone_repository/data/financial_ratios.csv")
```


