---
title: "Impute missing value"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(VIM)
library(tidyverse)
library(stringr)
library(data.table)
```


Import data that includes all columns from 3 statements
Proportion of missing value in each column 
```{r}
df <- read.csv('capstone_repository/data/df.csv')
```

Check the proportion of missing in tech sector
```{r}
col <-  c("Ticker","SimFinId", "Company.Name", "IndustryId","Sector", "Industry", "Report.Date","Fiscal.Year","Fiscal.Period", "Publish.Date", "Revenue", "Cost.of.Revenue", "Total.Assets","Total.Current.Assets","Total.Liabilities", "Total.Current.Liabilities", "Cash..Cash.Equivalents...Short.Term.Investments", "Accounts...Notes.Receivable", "Long.Term.Debt", "Total.Equity", "Short.Term.Debt", "Operating.Income..Loss.", "Interest.Expense..Net","Operating.Expenses", "Income.Tax..Expense..Benefit..Net", "Net.Income..Common.", "Shares..Diluted.", "Depreciation...Amortization")

df <- df[df$Sector == "Technology",col]
```

proportion of missing in each column
```{r}
sort(round(colMeans(!is.na(df))*100,2), decreasing = T)
```




Depreciation & Amortization
- 36 missing values
```{r}
#check number of missing values
sum(is.na(df$Depreciation...Amortization))
df$Depreciation...Amortization <- replace_na(df$Depreciation...Amortization, 0)
```



Long Term Debt
- 1015 missing values
```{r}
#check number of missing values
sum(is.na(df$Long.Term.Debt))

df$Long.Term.Debt <- replace_na(df$Long.Term.Debt, 0)
```



Income Tax Expense
- 42 missing values

```{r}
#check number of missing values
sum(is.na(df$Income.Tax..Expense..Benefit..Net))
df$Income.Tax..Expense..Benefit..Net <- replace_na(df$Income.Tax..Expense..Benefit..Net, 0)
```



Cost of Revenue
- 29 missing values
```{r}
#check number of missing values
sum(is.na(df$Cost.of.Revenue))
df$Cost.of.Revenue <- replace_na(df$Cost.of.Revenue, 0)
```


Operating Expense
- only 14 missing
```{r}
#check number of missing values
sum(is.na(df$Operating.Expenses))
df$Operating.Expenses <- replace_na(df$Operating.Expenses, 0)
```


Interest Expense, Net
- 1612 missing
```{r}
#check number of missing values
sum(is.na(df$Interest.Expense..Net))
df$Interest.Expense..Net <- replace_na(df$Interest.Expense..Net, 0)
```



Cash,Cash Equivalents & Short Term Investment
- 5 missing
```{r}
#check number of missing values
sum(is.na(df$Cash..Cash.Equivalents...Short.Term.Investments))
df$Cash..Cash.Equivalents...Short.Term.Investments <- replace_na(df$Cash..Cash.Equivalents...Short.Term.Investments, 0)
```



Accounts & Notes Receivable
- 30 missing
- on balance sheet, cash=0, thus reasonably conslude that null =0
```{r}
#check number of missing values
sum(is.na(df$Accounts...Notes.Receivable))
df$Accounts...Notes.Receivable <- replace_na(df$Accounts...Notes.Receivable, 0)

```


Short Term Debt
- 1267 missing
```{r}
#check number of missing values
sum(is.na(df$Short.Term.Debt))
df$Short.Term.Debt <- replace_na(df$Short.Term.Debt, 0)
```


```{r}
names(df)
cols <- c("Ticker","SimFinId", "Company.Name", "IndustryId","Sector", "Industry", "Report.Date","Fiscal.Year","Fiscal.Period", "Publish.Date", "Revenue", "Cost.of.Revenue", "Total.Assets","Total.Current.Assets","Total.Liabilities", "Total.Current.Liabilities", "Cash..Cash.Equivalents...Short.Term.Investments", "Accounts...Notes.Receivable", "Long.Term.Debt", "Total.Equity", "Short.Term.Debt", "Operating.Income..Loss.", "Interest.Expense..Net", "Depreciation...Amortization","Operating.Expenses", "Income.Tax..Expense..Benefit..Net", "Net.Income..Common.", "Shares..Diluted.")

allcol <- df[,cols]
colMeans(!is.na(allcol))

```



##########################################################计算ratio########################################################


Import libraries
```{r, warining = F}
library(tidyverse)
library(ggplot2)
library(VIM)
library(data.table)
```



Set up change id variables to character type, identify numeric columsn and categorical columns and logical columns
```{r}
ratio_tech <- copy(df)
```



(1)Current Ratio: Total Current Asset/ Total Current Liabilities 
```{r}
summary(ratio_tech$Total.Current.Assets)
sum(ratio_tech$Total.Current.Assets == 0) #1

summary(ratio_tech$Total.Current.Liabilities)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#0.000e+00 1.646e+08 6.461e+08 3.452e+09 2.324e+09 2.031e+11
sum(ratio_tech$Total.Current.Liabilities == 0) #1


#start calculation
ratio_tech <- ratio_tech %>% mutate(Current.Ratio = ifelse(Total.Current.Liabilities==0, NA, round(Total.Current.Assets/Total.Current.Liabilities,6)))

sum(is.na(ratio_tech$Current.Ratio)) #1
sum(ratio_tech$Current.Ratio == 0, na.rm = T)#0
```


(2)Quick Ratio: (Cash equivalent + Account receivable)/Total Current Liabilities
```{r}

summary(ratio_tech$Cash..Cash.Equivalents...Short.Term.Investments)
summary(ratio_tech$Accounts...Notes.Receivable)
summary(ratio_tech$Total.Current.Liabilities)

#start calculation
ratio_tech <- ratio_tech %>% mutate(Quick.Ratio = ifelse(Total.Current.Liabilities==0, NA, round((Cash..Cash.Equivalents...Short.Term.Investments + Accounts...Notes.Receivable)/Total.Current.Liabilities,6)))


sum(is.na(ratio_tech$Quick.Ratio)) #1
mean(ratio_tech$Quick.Ratio == 0, na.rm=T) #0

```

(3)Net working capital to asset ratio: (Total Current Asset - Total Current Liabilities)/ Total Current Asset
```{r}
#start calculation
ratio_tech <- ratio_tech %>% mutate(NWC.to.Asset.Ratio = ifelse(Total.Current.Assets==0, NA, round((Total.Current.Assets-Total.Current.Liabilities)/Total.Current.Assets,6)))

sum(is.na(ratio_tech$NWC.to.Asset.Ratio)) #1
sum(ratio_tech$NWC.to.Asset.Ratio == 0, na.rm = T) #0
sum(ratio_tech$NWC.to.Asset.Ratio < 0, na.rm = T) #225
```



(4)Cash Ratio: Cash Equivalent/ Total Current Liabilities
```{r}
#start calculation
ratio_tech <- ratio_tech %>% mutate(Cash.Ratio = ifelse(Total.Current.Liabilities==0, NA, round(Cash..Cash.Equivalents...Short.Term.Investments/Total.Current.Liabilities,6)))


sum(is.na(ratio_tech$Cash.Ratio))#1
sum(ratio_tech$Cash.Ratio == 0, na.rm = T) #6
```


(5)Long term Debt Ratio: Long term debt/(Long Term Debt + Total Equity)
we may want to drop this column
```{r}
#check if there are zeros or negative value that interrupt the calculation
summary(ratio_tech$Long.Term.Debt)
summary(ratio_tech$Total.Equity)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-4.632e+09  8.026e+07  3.854e+08  3.248e+09  1.584e+09  1.776e+11 
sum(ratio_tech$Total.Equity < 0, na.rm = T) #110


#start calculation
ratio_tech <- ratio_tech %>% mutate(Long.Term.Debt.Ratio = ifelse((Long.Term.Debt + Total.Equity)==0, NA, round(Long.Term.Debt/(Long.Term.Debt + Total.Equity),6)))


sum(ratio_tech$Long.Term.Debt.Ratio == 0, na.rm = T) #1110
sum(ratio_tech$Long.Term.Debt.Ratio < 0, na.rm = T) #24
```

(6)Liabilities to Equity Ratio: Total Liabilities/ Total Equity
```{r}
#check if there are zeros or negative value that interrupt the calculation
summary(ratio_tech$Total.Liabilities)
summary(ratio_tech$Total.Equity)

#start calculation
ratio_tech <- ratio_tech %>% mutate(Liabilities.to.Equity.Ratio = ifelse(Total.Equity==0, NA, round(Total.Liabilities/Total.Equity,6)))



#check result
#26 NA
#2 zero
#717 negative 
sum(is.na(ratio_tech$Liabilities.to.Equity.Ratio))#1
sum(ratio_tech$Liabilities.to.Equity.Ratio == 0, na.rm = T) #1
sum(ratio_tech$Liabilities.to.Equity.Ratio < 0, na.rm = T)#110
```



(7)Total Debt Ratio: Total Liabilities/Total Assets
```{r}

ratio_tech <- ratio_tech %>% mutate(Total.Debt.Ratio = ifelse(Total.Assets==0, NA, round(Total.Liabilities/Total.Assets,6)))


sum(is.na(ratio_tech$Total.Debt.Ratio))#0
sum(ratio_tech$Total.Debt.Ratio == 0, na.rm = T)#1
sum(ratio_tech$Total.Debt.Ratio < 0, na.rm = T)#0
```


(8)Debt to Asset Ratio:(Long term Debt + Short Term Debt)/ Total Assets
we may want to drop this
```{r}

#start calculation
ratio_tech <- ratio_tech %>% mutate(Debt.to.Asset.Ratio = ifelse(Total.Assets==0, NA, round((Long.Term.Debt + Short.Term.Debt)/Total.Assets,6)))

sum(is.na(ratio_tech$Debt.to.Asset.Ratio))#0
sum(ratio_tech$Debt.to.Asset.Ratio == 0, na.rm = T)#958
sum(ratio_tech$Debt.to.Asset.Ratio < 0, na.rm = T)#0
```


(9)Interest Coverage Ratio: Operating Income/Interest expense
this is too messy, we may want to discard it 
```{r}

summary(ratio_tech$Interest.Expense..Net)
sum(ratio_tech$Interest.Expense..Net <0)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-2.025e+09 -1.397e+07 -2.290e+05 -2.756e+07  0.000e+00  5.650e+08 

#总共分三种情况
#1. positive income & interst expense（正常）
sum(ratio_tech$Operating.Income..Loss. >0 & ratio_tech$Interest.Expense..Net <0) #1233
#2. positive income & interest gain （用mean fill）
sum(ratio_tech$Operating.Income..Loss. >0 & ratio_tech$Interest.Expense..Net >0) #360
#3. negative income & interest expense (正常)
sum(ratio_tech$Operating.Income..Loss. <0 & ratio_tech$Interest.Expense..Net <0) #409
#4. negative income & interest earn （不好， 异常 drop）
sum(ratio_tech$Operating.Income..Loss. <0 & ratio_tech$Interest.Expense..Net >0) #186



#start calculation
ratio_tech <- ratio_tech %>% mutate(Interest.Coverage.Ratio = ifelse(Interest.Expense..Net>=0, NA, round(Operating.Income..Loss./(-1*Interest.Expense..Net),6)))


#check result

sum(is.na(ratio_tech$Interest.Coverage.Ratio))#1205
sum(ratio_tech$Interest.Coverage.Ratio == 0, na.rm = T) #0
sum(ratio_tech$Interest.Coverage.Ratio < 0, na.rm = T)#409
```


(10)Cash Coverage Ratio:(Operating Income + Depreciation)/Interest expense
also kind of messy 
```{r}

#三种情况
#1. 分子大于 0 & interst expense（正常）
sum((ratio_tech$Depreciation...Amortization + ratio_tech$Operating.Income..Loss.) >0 & ratio_tech$Interest.Expense..Net < 0 ) #1365
#2. 分子大于 0 & interest gain （用mean fill）
sum((ratio_tech$Depreciation...Amortization + ratio_tech$Operating.Income..Loss.) >0 & ratio_tech$Interest.Expense..Net > 0 ) #391
#3. 分子小于 0 & interest expense (正常)
sum((ratio_tech$Depreciation...Amortization + ratio_tech$Operating.Income..Loss.) <0 & ratio_tech$Interest.Expense..Net < 0 ) #277
#4. 分子小于 0 & interest gain （不好， 异常 drop）
sum((ratio_tech$Depreciation...Amortization + ratio_tech$Operating.Income..Loss.) <0 & ratio_tech$Interest.Expense..Net > 0 ) #155


#start calculation
ratio_tech <- ratio_tech %>% mutate(Cash.Coverage.Ratio = ifelse(Interest.Expense..Net >= 0, NA, round((Operating.Income..Loss. + Depreciation...Amortization)/(-1*Interest.Expense..Net),6)))


sum(is.na(ratio_tech$Cash.Coverage.Ratio)) #1205
sum(ratio_tech$Cash.Coverage.Ratio == 0, na.rm = T)#0
sum(ratio_tech$Cash.Coverage.Ratio < 0, na.rm = T)#277
```



(11)After Tax Operating Income(Not a ratio)!!!: Revenue-Depreciaton-Income Tax expense
```{r}
#start calculation
summary(ratio_tech$Operating.Expenses)

ratio_tech <- ratio_tech %>% mutate(After.Tax.Operating.Income = ifelse(Operating.Expenses >0,NA,Revenue - Depreciation...Amortization + Operating.Expenses))

summary(ratio_tech$After.Tax.Operating.Income)
sum(ratio_tech$After.Tax.Operating.Income <0)#237
```


(12)Return On Assets: After Tax Operating Income/Total Assets
```{r}
#start calculation
ratio_tech <- ratio_tech %>% mutate(Return.On.Assets = After.Tax.Operating.Income/Total.Assets)

sum(is.na(ratio_tech$Return.On.Assets))#0
```



(13)Return On Capital: After Tax Operating Income/(Long Term Debt + Total Equity)
```{r}
summary(ratio_tech$Total.Equity)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.453e+10  1.356e+08  7.464e+08  4.045e+09  2.894e+09  3.525e+11 
sum(ratio_tech$Total.Equity <0 , na.rm = T)#110


#4钟情况
#1.分子正 & 分母正 （正常）
sum(ratio_tech$After.Tax.Operating.Income >0 & (ratio_tech$Long.Term.Debt + ratio_tech$Total.Equity) > 0) #2571
#2.分子正 & 分母负 （drop）21个
sum(ratio_tech$After.Tax.Operating.Income >0 & (ratio_tech$Long.Term.Debt + ratio_tech$Total.Equity) < 0) #38
#3.分子负 & 分母正 （正常）
sum(ratio_tech$After.Tax.Operating.Income <0 & (ratio_tech$Long.Term.Debt + ratio_tech$Total.Equity) > 0) #223
#4.分子负 & 分母负 （drop）5个
sum(ratio_tech$After.Tax.Operating.Income <0 & (ratio_tech$Long.Term.Debt + ratio_tech$Total.Equity) < 0) #14


#start calculation
ratio_tech <- ratio_tech %>% mutate(Return.On.Capital = ifelse(((Long.Term.Debt + Total.Equity) <= 0),NA,After.Tax.Operating.Income/(Long.Term.Debt + Total.Equity)))

sum(is.na(ratio_tech$Return.On.Capital)) #52
sum(ratio_tech$Return.On.Capital <0, na.rm=T) #223
```



(14)Return On Equity: After Tax Operating Income/Total Equity
```{r}

summary(ratio_tech$Total.Equity)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-1.453e+10  1.356e+08  7.464e+08  4.045e+09  2.894e+09  3.525e+11 
sum(ratio_tech$Total.Equity < 0 , na.rm = T)#110


sum(ratio$Total.Equity < 0 & ratio$After.Tax.Operating.Income < 0 & ratio$Sector == 'Technology') 
#4钟情况
#1.分子正 & 分母正 （正常）
#2.分子正 & 分母负 （drop）86个
#3.分子负 & 分母正 （正常）
#4.分子负 & 分母负 （drop）24个

#start calculation
#we need to create a indicator telling whether 
ratio_tech <- ratio_tech %>% mutate(Return.On.Equity = ifelse( Total.Equity <= 0,NA,After.Tax.Operating.Income/Total.Equity))


sum(is.na(ratio_tech$Return.On.Equity)) #111
sum(ratio_tech$Return.On.Equity <0, na.rm=T) #214
```



(15)Operating Profit Margin: Operating Income/ Revenue
```{r}

#start calculation
ratio_tech <- ratio_tech %>% mutate(Operating.Profit.Margin = ifelse( Revenue == 0,NA,Operating.Income..Loss./Revenue))
sum(is.na(ratio_tech$Operating.Profit.Margin)) #6
sum(ratio_tech$Operating.Profit.Margin <0, na.rm=T) #787
```



(16)Net Profit Margin: Net Income/Revenue

```{r}


#start calculation
ratio_tech <- ratio_tech %>% mutate(Net.Profit.Margin = ifelse( Revenue == 0,NA,Net.Income..Common./Revenue))
sum(is.na(ratio_tech$Net.Profit.Margin)) #6
sum(ratio_tech$Net.Profit.Margin <0, na.rm=T) #939
```

(17)Asset Turnover; Revenue/Total Asset(year-start)
```{r, message warning}
#create a variable of year-start total asset
ratio_tech <- as.data.table(ratio_tech)
ratio_tech[, Total.Assets.year.start := shift(Total.Assets, 1, 'lag'), by = Ticker]

sum(is.na(ratio_tech$Total.Assets.year.start)) # 377 missing


#calculate ratio
ratio_tech <- ratio_tech %>% mutate(Asset.Turnover = ifelse( Total.Assets.year.start == 0,NA,Revenue/Total.Assets.year.start))
sum(is.na(ratio_tech$Asset.Turnover)) #377
sum(ratio_tech$Asset.Turnover <0, na.rm=T) #0
```



(18)Receivable Turnover: Revenue/Account receivable year start
```{r, message warning}
#create a variable of year-start account receivable
ratio_tech <- as.data.table(ratio_tech)
ratio_tech[, Accounts...Notes.Receivable.year.start := shift(Accounts...Notes.Receivable, 1, 'lag'), by = Ticker]

sum(is.na(ratio_tech$Accounts...Notes.Receivable.year.start)) # 377 missing
sum(ratio_tech$Accounts...Notes.Receivable.year.start ==0, na.rm = T) #28 zero


#calculate ratio
ratio_tech <- ratio_tech %>% mutate(Receivable.Turnover = ifelse(Accounts...Notes.Receivable.year.start == 0, NA, Revenue/Accounts...Notes.Receivable.year.start))

sum(is.na(ratio_tech$Receivable.Turnover)) #405
sum(ratio_tech$Receivable.Turnover <0, na.rm=T) #0
```




(19)Book to Market Ratio: Total Equity /Stcok Price(year-end) * Shares(basic)
```{r}
#import stock price data
stock <- fread('simfin_data/us-shareprices-daily.csv')
stock$Year <- str_split_fixed(stock$Date,"-", n=3)[,1]
stock$Date2 <- str_remove_all(stock$Date, "-")


#find the year-end stock price 
end_price <- stock %>% group_by(Ticker, Year) %>% mutate(year_end = max(Date2)) %>% filter(Date2 == year_end)%>% select(Ticker, SimFinId, Year, Close)
end_price$SimFinId <- as.character(end_price$SimFinId)


#join it with statement data 
ratio_tech$Ticker <- as.character(ratio_tech$Ticker)
ratio_tech$SimFinId <- as.character(ratio_tech$SimFinId)
ratio_tech$Fiscal.Year <- as.character(ratio_tech$Fiscal.Year)
ratio_tech <- left_join(ratio_tech,end_price, by= c("Ticker", "SimFinId", "Fiscal.Year" = "Year"))
sum(is.na(ratio_tech$Close)) #256


#start calculation 
ratio_tech <- ratio_tech %>% mutate(Book.to.Market.Ratio = ifelse(Close*Shares..Diluted. ==0 , NA, (Total.Equity/(Shares..Diluted. * Close))))

sum(is.na(ratio_tech$Book.to.Market.Ratio)) #256
sum(ratio_tech$Book.to.Market.Ratio <0, na.rm=T) #75
```

(20)Earning Per Share: Net Income/ Shares
```{r}

#start calculation 
ratio_tech <- ratio_tech %>% mutate(Earning.Per.Share = ifelse(Shares..Diluted. ==0 , NA, Net.Income..Common./Shares..Diluted.))

sum(is.na(ratio_tech$Earning.Per.Share)) #0
sum(ratio_tech$Earning.Per.Share <0, na.rm=T) #943
summary(ratio_tech$Earning.Per.Share)
```


(21)Sales Per Share: Revenue/ Shares
```{r}

#start calculation 
ratio_tech <- ratio_tech %>% mutate(Sales.Per.Share = ifelse(Shares..Diluted. ==0 , NA, Revenue/Shares..Diluted.))

sum(is.na(ratio_tech$Sales.Per.Share)) #0
sum(ratio_tech$Sales.Per.Share <0, na.rm=T) #0
```


(22)Pricing to Earning Ratio: Stock Price/ Earning Per Share
```{r}
#start calculation 
ratio_tech <- ratio_tech %>% mutate(Price.to.Earning.Ratio = ifelse(Earning.Per.Share ==0 , NA, Close/Earning.Per.Share))

sum(is.na(ratio_tech$Price.to.Earning.Ratio)) #257
sum(ratio_tech$Price.to.Earning.Ratio <0, na.rm=T) #791
```



######################################################Select Financial ratios needed #################################################


Select ratios needed 
```{r}
names(ratio_tech)
tech2 <- ratio_tech[,c(1:8, 29:32, 34,35, 40:44, 46, 48, 50:53)]
names(tech2)

dim(na.omit(tech2)) #2233 25

```


highly correlated columsn
1.  Current Ratio & Quick Ratio & Cash Ratio 
2.  Return on Equity & Liabilities to Equity Ratio
3.  Net.Profit.Margin  & Operating Profit Margin
4.  Receivable.Turnover & Assets Turnover
5.  Earning.Per.Share & Book.to Market Ratio & Sales.Per.Share
```{r}
#cor(na.omit(tech2[-c(1:8)]))
```


####################################################Join with response variable########################################



```{r}
library(dplyr)
library(tidyverse)

#Set up
stock <- read.csv2("simfin_data/us-shareprices-daily.csv")
stock2 <- stock[c(1,2,3,4,7,9)] ## select relative columns

stock2$Year <- substring(date,1,4) ## extract Year
stock2$Date <- str_remove_all(stock2$Date, "-")
stock2$day <- str_sub(str_remove_all(stock2$Date, "-"), start = 5)

date <- as.Date(stock2$Date)


## fill NAs values in Dividend with 0
stock2$Dividend <- as.numeric(as.character(stock2$Dividend))
stock2$Dividend <- replace_na(stock2$Dividend,0)
```





```{r}
#get the sum of dividends
stock2$day <- as.integer(stock2$day)
stock2$Year <- as.integer(stock2$Year)
stock2 <- stock2 %>% group_by(Year) %>% mutate(Year.adj = ifelse(day <= 1031, Year, Year+1))
totDiv.adj <- stock2 %>% group_by(Ticker, Year.adj) %>% summarise(totDiv.adj =sum(Dividend)) ##total Dividend within a year


#get the stock price on 10-31 each year
first_last2 <- stock2 %>% group_by(Ticker, Year.adj) %>% filter(row_number() %in% c(1, n()))
first_last2$Open <- as.numeric(as.character(first_last2$Open))
first_last2$Close <- as.numeric(as.character(first_last2$Close))


#calculate stock return 

stock_return <- first_last2[!duplicated(first_last2[c("Ticker","Year.adj")]), c("Ticker","SimFinId","Year.adj")]
totDiv.adj$Year.adj <- as.character(totDiv.adj$Year.adj)
stock_return$Year.adj <- as.character(stock_return$Year.adj)
stock_return <- left_join(stock_return, totDiv.adj, by=c("Ticker" , "Year.adj")) ## merge annual Dividend

stock_return$Return <- 0
first_last2$Year.adj <- as.character(first_last2$Year.adj)
first_last2$Ticker <- as.character(first_last2$Ticker)
stock_return$Ticker <- as.character(stock_return$Ticker)
for (i in 1:nrow(stock_return)) {
  thisTicker <- stock_return$Ticker[i]
  thisYear <- stock_return$Year.adj[i]
  open <- as.numeric(first_last2[first_last2$Ticker==thisTicker & first_last2$Year.adj==thisYear, "Open"][1,])
  close <- as.numeric(first_last2[first_last2$Ticker==thisTicker & first_last2$Year.adj==thisYear, "Close"][2,])
  div <- as.numeric(stock_return$totDiv.adj[i])
  stock_return$Return[i] <- (log(close + div) - log(open))
}
dim(stock_return)

stock_return <- stock_return[c("Ticker", "Year.adj", "Return")]
```



join with financial ratios
```{r}
stock_return$Year.adj.adj <-  as.character(as.integer(stock_return$Year.adj) -1L)

xy <- left_join(tech2, stock_return, by = c("Ticker", "Fiscal.Year" = "Year.adj")) #t和tjoin 
xy <- left_join(xy, stock_return, by =c("Ticker", "Fiscal.Year" = "Year.adj.adj")) #return_t 和ratio_t-1 join

#get rid of unnecessary columsn
xy$Year.adj <- NULL
xy$Year.adj.adj <- NULL
```


####################################################Variables Transformation###########################################


highly correlated columsn
1.  Current Ratio & Quick Ratio & Cash Ratio
2.  Return on Equity & Liabilities to Equity Ratio
3.  Net.Profit.Margin  & Operating Profit Margin
4.  Receivable.Turnover & Assets Turnover
5.  Earning.Per.Share & Sales.Per.Share
```{r}
cor(na.omit(xy[-c(1:8)]))
```


Compare Current Ratio & Quick Ratio & Cash Ratio
- All pretty symmetric after log transformation
- Current ratio has the highest correlation, so keep it
```{r}
hist(Math.log(na.omit(xy$Current.Ratio)))
hist(Math.log(na.omit(xy$Cash.Ratio)))
hist(Math.log(na.omit(xy$Quick.Ratio)))

Math.log <- function(x){
   min_x <- min(x, na.rm = T)
   log(x + abs(min_x) + 0.1) }

xy$Current.Ratio.log <- Math.log(xy$Current.Ratio)
xy$Cash.Ratio.log <- Math.log(xy$Cash.Ratio)
xy$Quick.Ratio.log <- Math.log(xy$Quick.Ratio)


cor(na.omit(xy[c("Current.Ratio","Current.Ratio.log", "Cash.Ratio.log", "Quick.Ratio.log", "Return.x", "Return.y")]))
```

```{r}
xy_t <- copy(xy)

xy_t$Cash.Ratio <- NULL
xy_t$Cash.Ratio.log <- NULL
xy_t$Quick.Ratio <- NULL
xy_t$Quick.Ratio.log <- NULL
xy_t$Current.Ratio <- NULL
```



Compare Return on Equity & Liabilities to Equity Ratio
-Liabilities.to.Equity has fewer mising value but much more outliers, and its correlation with return is also lower 
-we are keeping the Return on Equity with log transformation
```{r}
summary(xy$Return.On.Equity)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max.       NA's 
#-19651.202      0.297      0.594     -5.412      1.248    964.798        111 
summary(xy$Liabilities.to.Equity.Ratio)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max.       NA's 
#-1354.0182     0.3046     0.6312     0.5290     1.3308   917.1449          1 

#check number of outliers
sum(xy$Return.On.Equity < -3, na.rm = T) #10
sum(xy$Liabilities.to.Equity.Ratio < -3, na.rm = T) #99


#use cube root for transforming return on equity
hist(Math.cbrt(na.omit(xy$Return.On.Equity[xy$Return.On.Equity > -2 & xy$Return.On.Equity <2] )))
hist(Math.log(na.omit(xy$Return.On.Equity[xy$Return.On.Equity > -3] )))
hist(Math.cbrt(na.omit(xy$Liabilities.to.Equity.Ratio[xy$Liabilities.to.Equity.Ratio > -3])))


Math.log <- function(x){
   min_x <- min(x, na.rm = T)
   log(x + abs(min_x) + 0.1) 
   }

xy$Return.On.Equity.log <- Math.log(xy$Return.On.Equity)
xy$Return.On.Equity.cube <- Math.cbrt(xy$Return.On.Equity)
xy$Liabilities.to.Equity.Ratio.log <- Math.log(xy$Liabilities.to.Equity.Ratio)
xy$Liabilities.to.Equity.Ratio.cube <- Math.cbrt(xy$Liabilities.to.Equity.Ratio)


cor(na.omit(xy[c("Return.On.Equity","Return.On.Equity.log", "Return.On.Equity.cube", "Liabilities.to.Equity.Ratio", "Liabilities.to.Equity.Ratio.log","Liabilities.to.Equity.Ratio.cube", "Return.x", "Return.y")]))
```


```{r}
xy_t$Return.On.Equity <- xy$Return.On.Equity
xy_t$Return.On.Equity.log <- Math.log(xy_t$Return.On.Equity)

```


```{r}
xy_t$Liabilities.to.Equity.Ratio <- NULL

```



Compare Net.Profit.Margin  & Operating Profit Margin
- based on correlation keep Operating Profit Margin with cube transformation
```{r}
hist(Math.log(na.omit(xy$Net.Profit.Margin)))
hist(Math.log(na.omit(xy$Operating.Profit.Margin)))

xy$Net.Profit.Margin.log <- Math.log(xy$Net.Profit.Margin)
xy$Operating.Profit.Margin.log <- Math.log(xy$Operating.Profit.Margin)
xy$Net.Profit.Margin.cube <- Math.cbrt(xy$Net.Profit.Margin)
xy$Operating.Profit.Margin.cube <- Math.cbrt(xy$Operating.Profit.Margin)


cor(na.omit(xy[c("Net.Profit.Margin","Net.Profit.Margin.cube","Net.Profit.Margin.log", "Operating.Profit.Margin","Operating.Profit.Margin.cube", "Operating.Profit.Margin.log", "Return.x", "Return.y")]))
```


```{r}
xy_t$Net.Profit.Margin <- NULL
xy_t$Operating.Profit.Margin <- NULL
xy_t$Operating.Profit.Margin.cube <- xy$Operating.Profit.Margin.cube

```



Compare Receivable.Turnover & Assets Turnover
after logging, the correlation between them are not so high  keep them both
```{r}
summary(xy$Receivable.Turnover)
summary(xy$Asset.Turnover)

hist(Math.log(na.omit(xy$Receivable.Turnover)))
hist(Math.log(na.omit(xy$Asset.Turnover)))

xy$Receivable.Turnover.log <- Math.log(xy$Receivable.Turnover)
xy$Asset.Turnover.log <- Math.log(xy$Asset.Turnover)


cor(na.omit(xy[c("Receivable.Turnover","Receivable.Turnover.log", "Asset.Turnover", "Asset.Turnover.log", "Return.x", "Return.y")]))
```


```{r}
xy_t$Receivable.Turnover.log <- xy$Receivable.Turnover.log
xy_t$Asset.Turnover.log <- xy$Asset.Turnover.log
xy_t$Receivable.Turnover <- NULL
xy_t$Asset.Turnover <- NULL
```




Compare  Earning.Per.Share  & Sales.Per.Share
- keep earnin per share after log transformation
```{r}
summary(xy$Earning.Per.Share)
summary(xy$Sales.Per.Share)
hist(Math.log(na.omit(xy$Earning.Per.Share)))
hist(Math.log(na.omit(xy$Sales.Per.Share)))

xy$Earning.Per.Share.log <- Math.log(xy$Earning.Per.Share)
xy$Sales.Per.Share.log <- Math.log(xy$Sales.Per.Share)



cor(na.omit(xy[c("Earning.Per.Share","Earning.Per.Share.log", "Sales.Per.Share", "Sales.Per.Share.log", "Return.x", "Return.y")]))
```

```{r}
xy_t$Earning.Per.Share <- Math.log(xy_t$Earning.Per.Share)
xy_t$Earning.Per.Share <- NULL
xy_t$Sales.Per.Share <- NULL


```

"NWC.to.Asset.Ratio" 
```{r}
summary(xy$NWC.to.Asset.Ratio)

hist(na.omit(xy$NWC.to.Asset.Ratio))

hist(Math.log(na.omit(xy$NWC.to.Asset.Ratio)))

xy$NWC.to.Asset.Ratio.log <- Math.log(xy$NWC.to.Asset.Ratio)
xy$NWC.to.Asset.Ratio.cube <- Math.cbrt(xy$NWC.to.Asset.Ratio)

cor(na.omit(xy[c("NWC.to.Asset.Ratio","NWC.to.Asset.Ratio.log","NWC.to.Asset.Ratio.cube", "Return.x", "Return.y")]))

xy_t$NWC.to.Asset.Ratio.log <- Math.log(xy_t$NWC.to.Asset.Ratio)
xy_t$NWC.to.Asset.Ratio <- NULL
```


Total.Debt.Ratio

```{r}
summary(xy$Total.Debt.Ratio)

hist(na.omit(xy$Total.Debt.Ratio))

hist(Math.log(na.omit(xy$Total.Debt.Ratio)))

xy$Total.Debt.Ratio.log <- Math.log(xy$Total.Debt.Ratio)
xy$NWC.to.Asset.Ratio.cube <- Math.cbrt(xy$NWC.to.Asset.Ratio)

cor(na.omit(xy[c("Total.Debt.Ratio","Total.Debt.Ratio.log", "Return.x", "Return.y")]))

xy_t$Total.Debt.Ratio.log <- Math.log(xy_t$Total.Debt.Ratio)
xy_t$Total.Debt.Ratio <- NULL
```


"Return.On.Assets" 

```{r}
summary(xy$Return.On.Assets)
sum(xy$Return.On.Assets < -3) #2

hist(na.omit(xy$Return.On.Assets[xy$Return.On.Assets > -3]))

hist(Math.log(na.omit(xy$Return.On.Assets[xy$Return.On.Assets > -3])))

#remove outliers 
xy$Return.On.Assets[xy$Return.On.Assets < -3] <- NA
xy$Return.On.Assets.log <- Math.log(xy$Return.On.Assets)

cor(na.omit(xy[c("Return.On.Assets","Return.On.Assets.log", "Return.x", "Return.y")]))

xy_t$Return.On.Assets.log <- Math.log(xy_t$Return.On.Assets)
xy_t$Return.On.Assets <- NULL
```


"Return.On.Capital" 
- highly correlated with return on asset and has more messy distribution
```{r}
xy_t$Return.On.Capital <- NULL
```


"Book.to.Market.Ratio"     

```{r}
summary(xy$Book.to.Market.Ratio)
sum(!xy$Book.to.Market.Ratio %between% c(-10,10),na.rm = T) #17

hist(na.omit(xy$Book.to.Market.Ratio[xy$Book.to.Market.Ratio %between% c(-10,10)]))

hist(Math.log(na.omit(xy$Book.to.Market.Ratio[xy$Book.to.Market.Ratio %between% c(-10,10)])))

#remove outliers 
xy$Book.to.Market.Ratio[!xy$Book.to.Market.Ratio %between% c(-10,10)] <- NA
xy$Book.to.Market.Ratio.log <- Math.log(xy$Book.to.Market.Ratio)

cor(na.omit(xy[c("Book.to.Market.Ratio","Book.to.Market.Ratio.log", "Return.x", "Return.y")]))

xy_t$Book.to.Market.Ratio[!xy_t$Book.to.Market.Ratio %between% c(-10,10)] <- NA
xy_t$Book.to.Market.Ratio.log <- Math.log(xy_t$Book.to.Market.Ratio)
```


"Price.to.Earning.Ratio"


```{r}
summary(xy$Price.to.Earning.Ratio)
sum(!xy$Price.to.Earning.Ratio %between% c(-300,300),na.rm = T) #87

hist(na.omit(xy$Price.to.Earning.Ratio[xy$Price.to.Earning.Ratio %between% c(-300,300)]))

hist(Math.log(na.omit(xy$Price.to.Earning.Ratio[xy$Price.to.Earning.Ratio %between% c(-300,300)])))

#remove outliers 
xy$Price.to.Earning.Ratio[!xy$Price.to.Earning.Ratio %between% c(-300,300)] <- NA
xy$Price.to.Earning.Ratio.log <- Math.log(xy$Price.to.Earning.Ratio)

cor(na.omit(xy[c("Price.to.Earning.Ratio","Price.to.Earning.Ratio.log", "Return.x", "Return.y")]))

xy_t$Price.to.Earning.Ratio[!xy_t$Price.to.Earning.Ratio %between% c(-300,300)] <- NA
xy_t$Price.to.Earning.Ratio.log <- Math.log(xy_t$Price.to.Earning.Ratio)

```


```{r}
test <- na.omit(xy_t)

cor(test[-c(1:8)])
```


```{r}
summary(glm(Return.y ~ . , data = test[-c(1:8, 12)]))
```


```{r}
summary(glm(Return.y ~ . , data = xy[-c(1:8)]))
```

############################################################start over again####################################################
Finally we picked 
1. Operating profit margin
2. receivable turn over 
3. Total Debt ratio
4. Saales Per share 
5. Book to market ratio
6. Return on Equity 

```{r}
xy$Fiscal.Year <- as.integer(xy$Fiscal.Year)
summary(lm(Return.y ~ Operating.Profit.Margin + Receivable.Turnover + Total.Debt.Ratio + Sales.Per.Share + Book.to.Market.Ratio + Return.On.Equity, data = xy[xy$Fiscal.Year < 2017,]))
```


```{r}

model <- lm(Return.y ~ Operating.Profit.Margin + Receivable.Turnover + Total.Debt.Ratio + Sales.Per.Share + Book.to.Market.Ratio + Return.On.Equity, data = xy[xy$Fiscal.Year < 2017,])

test2018 <- xy[xy$Fiscal.Year == 2018,]

distPred <- predict(model, test2018)  # predict distance
actuals_preds <- data.frame(cbind(test2018[1:8], actuals=test2018$Return.y, predicteds=distPred)) 

#good
actuals_preds[order(-actuals_preds$predicteds),][1:10,] %>% summarise(totreturn = sum(predicteds))
actuals_preds[order(actuals_preds$predicteds),][1:10,] %>% summarise(totreturn = sum(predicteds))
```

