---
title: "Exploratory Log Transformation of Ratios"
author: "Zhenzhen Zhu (zz8vy)"
date: "4/4/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(MASS)
library(tidyverse)
library(data.table)
#df <- fread("allcol.csv", drop = 1)
#stock <- read.csv2("us-shareprices-daily.csv")
ratio_return <- read.csv("ratios_return.csv", header = T) ## 11718 * 32
ratio_return$Sector <- as.character(ratio_return$Sector)
tech <- ratio_return[ratio_return$Sector=="Technology",] ## 2085 observations
length(levels(tech$Ticker)) ##1584 companies
```


## Return on Assets
```{r}
summary(tech$Return.On.Assets)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#-3.7872  0.1932  0.3774  0.5030  0.6718  4.3005 
plot(tech$Return.On.Assets)
plot(density(tech$Return.On.Assets))
## translate (add minimum negative value), then log transform
log_roa <- log(tech$Return.On.Assets+1+abs(min(tech$Return.On.Assets)))
tech$Return.On.Assets.t <- log_roa
plot(log_roa)
plot(density(log_roa)) ## still left-skewed
plot(density(tech$Return.On.Assets^1/3))

## outliers:
tech[log_roa<1, c("Ticker", "Fiscal.Year")]
# log_roa[which(log_roa<1)] ## 0.7893779 0.0000000 0.7815722 1.2369020 1.2719327
# plot(density(log_roa[log_roa>=1]))

### outliers lying >3*SD away of mean
sd <- sd(tech$Return.On.Assets) 
mean <- mean(tech$Return.On.Assets) 
ci <- c(mean - 3*sd, mean + 3*sd) 
#sum(!tech$Return.On.Assets %between% ci)#45 outliers
tech[!tech$Return.On.Assets %between% ci, c('Ticker', 'Fiscal.Year')] ##outliers
tech$Return.On.Assets.out <- "False"
tech$Return.On.Assets.out[which(!tech$Return.On.Assets %between% ci)] <- "True"
plot(density(tech$Return.On.Assets.t[tech$Return.On.Assets.out=="False"]))


## correlation with Stock Return
cor(log_roa, tech$Return)
cor(log_roa, tech$Return2)
cor(log_roa, tech$Return3) #0.04531882
cor(tech$Return.On.Assets, tech$Return3) #0.04179654
cor(tech$Return.On.Assets^1/3, tech$Return3) #0.04179654
cor(log_roa, tech$Return4)
```
Distribution doesn't change much after taking log transformation or cube root.
Suggestion: no need to tansform Return.On.Assets. If you are to transform, log transformation is better than cube root because it has highest correlation with Return3.

Outliers: 
   Ticker   Fiscal.Year)
	MRAM	2016		
	MRAM	2017		
	MRAM	2018	
Note: This company (EVERSPIN TECHNOLOGIES INC) only has three years in the data set, and all three years are outliers. 



## Earning per Share
```{r}
summary(tech$Earning.Per.Share)
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
# -21.9880   -0.0395    0.6528    2.4637    1.8690 1260.5818
sum(tech$Earning.Per.Share<0) ## 710 (==0: 0)
plot(tech$Earning.Per.Share)
plot(density(tech$Earning.Per.Share)) ## severely right-skewed

## transformamtion
log_eps <- log(tech$Earning.Per.Share+1+abs(min(tech$Earning.Per.Share)))
tech$Earning.Per.Share.t <- log_eps
plot(log_eps)
plot(density(log_eps))
plot(density(tech$Earning.Per.Share^1/3))

## outliers
tech[(log_eps<1) | (log_eps>5), c("Ticker", "Fiscal.Year")]
plot(density(log_eps[log_eps<=5 & log_eps>=1]))
tech$Earning.Per.Share.out <- "False"
tech$Earning.Per.Share.out[which(log_eps<1) | (log_eps>5)] <- "True"

### Note: The ">3*SD" rule doesn't work because the points within the 3*SD are still left skewed,
### hence, remove points where log transformed values are <1 or >5

sd <- sd(tech$Earning.Per.Share) 
mean <- mean(tech$Earning.Per.Share) 
ci <- c(mean - 3*sd, mean + 3*sd) 
#sum(!tech$Earning.Per.Share %between% ci)# 4 outliers
tech[!tech$Earning.Per.Share %between% ci, c('Ticker', 'Fiscal.Year')] ##outliers
tech$Earning.Per.Share.out <- "False"
tech$Earning.Per.Share.out[which(!tech$Earning.Per.Share %between% ci)] <- "True"
plot(density(tech$Earning.Per.Share.t[tech$Earning.Per.Share.out=="False"])) ## still left skewed


cor(log_eps, tech$Return)
cor(log_eps, tech$Return2)
cor(log_eps, tech$Return3) #0.1226962
cor(tech$Earning.Per.Share, tech$Return3) 
cor(tech$Earning.Per.Share^1/3, tech$Return3) 
cor(log_eps, tech$Return4)
```
Original Earning per Share is heavily right-skewed. Log transformation will fix this. It also leads to highest correlation with Return3.

Outliers (rule: log Earning per Share <1 or >5)
Ticker   Year
HURC	2013		(highest Earning per Share, >7)
IEHC	2015		
IEHC	2016		(between 6 and 7)
IPDN	2015		(a little above 5)
RWC	2012    (lowest, 0)



## Sales per Share
```{r}
summary(tech$Sales.Per.Share) 
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#    0.076     5.444     9.449    47.031    18.639 29675.850
plot(tech$Sales.Per.Share)
plot(density(tech$Sales.Per.Share))

## transformamtion
log_sps <- log(tech$Sales.Per.Share)
tech$Sales.Per.Share.t <- log_sps
plot(log_sps)
plot(density(log_sps))

## outliers
fiveLargest <- tail(order(tech$Sales.Per.Share), 5) ## indice for the five highest sales per share
tech[fiveLargest, c("Ticker", "Fiscal.Year")]

### outliers lying >3*SD away of mean
sd <- sd(tech$Sales.Per.Share) 
mean <- mean(tech$Sales.Per.Share) 
ci <- c(mean - 3*sd, mean + 3*sd)
#sum(!tech$Sales.Per.Share %between% ci)#4 outliers
tech[!tech$Sales.Per.Share %between% ci, c('Ticker', 'Fiscal.Year')] ##outliers
tech$Sales.Per.Share.out <- "False"
tech$Sales.Per.Share.out[which(!tech$Sales.Per.Share %between% ci)] <- "True"
plot(density(tech$Sales.Per.Share.t[tech$Sales.Per.Share.out=="False"]))


cor(log_sps, tech$Return)
cor(log_sps, tech$Return2)
cor(log_sps, tech$Return3) #0.02757461
cor(tech$Sales.Per.Share, tech$Return3)
cor(log_sps, tech$Return4)
```
The original data is too dense around small numbers close to 0, with a few extreme large outliers.
I suggest taking log transformation

Outliers (5 largest, sorted descendingly in terms of Sales per Share)
HURC	2013
IEHC	2016
IEHC	2015
RWC	2018
RWC	2012		

Note: HURC 2013, IEHC 2015, IEHC 2016, RWC 2012 are also outliers for Earning per Share
This implies that these companies performed well in these years.



## Price to Earning Ratio
```{r}
summary(tech$Price.to.Earning.Ratio)
#      Min.    1st Qu.     Median       Mean    3rd Qu.       Max. 
#-11489.062     -1.291     18.765     26.740     35.296   9682.311  
sum(tech$Price.to.Earning.Ratio<0) ## 710 (==0: 0)
plot(tech$Price.to.Earning.Ratio)
plot(density(tech$Price.to.Earning.Ratio))

## transformamtion
log_per <- log(tech$Price.to.Earning.Ratio+1+abs(min(tech$Price.to.Earning.Ratio)))
tech$Price.to.Earning.Ratio.t <- log_per
plot(log_per)
plot(density(log_per))

## outliers
x <- which(tech$Price.to.Earning.Ratio == min(tech$Price.to.Earning.Ratio) |
              tech$Price.to.Earning.Ratio == max(tech$Price.to.Earning.Ratio))
tech[x, c("Ticker", "Fiscal.Year")]

## reperform transformamtion after removing FIVN 2018
log_per[x] <- NA
plot(log_per)
plot(density(log_per[!is.na(log_per)]))

cor(log_per, tech$Return, use="complete.obs")
cor(log_per, tech$Return2, use="complete.obs")
cor(log_per, tech$Return3, use="complete.obs") #0.05069594
cor(tech$Price.to.Earning.Ratio, tech$Return3)
cor(log_per, tech$Return4, use="complete.obs")



### outliers lying >3*SD away of mean
sd <- sd(tech$Price.to.Earning.Ratio) 
mean <- mean(tech$Price.to.Earning.Ratio) 
ci <- c(mean - 3*sd, mean + 3*sd)
#sum(!tech$Price.to.Earning.Ratio %between% ci)# 17 outliers
tech[!tech$Price.to.Earning.Ratio %between% ci, c('Ticker', 'Fiscal.Year')] ##outliers
tech$Price.to.Earning.Ratio.out <- "False"
tech$Price.to.Earning.Ratio.out[which(!tech$Price.to.Earning.Ratio %between% ci)] <- "True"
plot(density(tech$Price.to.Earning.Ratio.t[tech$Price.to.Earning.Ratio.out=="False"]))


```

Outliers:
FIVN	2018	(lowest)
RP	2017	(highest)

Suggestion: Remove FIVN	2018 since it's incredibly lower than other observations, even after taking the log transformaion.

Also may consider removing the highest Price to Earning Ratio, which is RP	2017.
After taking out the lowest and the highest, and performing log transformation, the density curve is a little right-skewed. To solve this, you may continue to remove more outliers with log_per >=9.6 (there are three of them)



## Book to Market Ratio
```{r}
summary(tech$Book.to.Market.Ratio)
# Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#   0.0042    0.1965    0.3437    2.8135    0.6151 1326.5407
plot(tech$Book.to.Market.Ratio) ## a few outliers with pretty high Book to Market Ratio

## transformamtion
log_bmr <- log(tech$Book.to.Market.Ratio)
tech$Book.to.Market.Ratio.t <- log_bmr
plot(log_bmr)
plot(density(log_bmr)) ## a little right-skewed

## outliers
fiveLargest <- tail(order(tech$Book.to.Market.Ratio), 5) ## indice for the five highest sales per share
tech[fiveLargest, c("Ticker", "Fiscal.Year")]

### outliers lying >3*SD away of mean
sd <- sd(tech$Book.to.Market.Ratio) 
mean <- mean(tech$Book.to.Market.Ratio) 
ci <- c(mean - 3*sd, mean + 3*sd)
#sum(!tech$Book.to.Market.Ratio %between% ci)# 5 outliers
tech[!tech$Book.to.Market.Ratio %between% ci, c('Ticker', 'Fiscal.Year')] ##outliers
tech$Book.to.Market.Ratio.out <- "False"
tech$Book.to.Market.Ratio.out[which(!tech$Book.to.Market.Ratio %between% ci)] <- "True"
plot(density(tech$Book.to.Market.Ratio.t[tech$Book.to.Market.Ratio.out=="False"]))



cor(log_bmr, tech$Return)
cor(log_bmr, tech$Return2)
cor(log_bmr, tech$Return3) #-0.2381094
cor(tech$Price.to.Earning.Ratio, tech$Return3) #0.02814808
cor(log_bmr, tech$Return4)
```
Note: The original Book to Market Ratio has slightly higher correlation with Return3 than log transformed data.
Log transformed data is right-skewed.


Outliers:
RWC	2012
IEHC	2016
HURC	2013
IEHC	2015
RWC	2018		

Note: Same outliers as Sales per Share. Mostly the same as Earning per Share.



## write columns to .csv file
```{r}
ratio_transformed_zhenzhen <- tech[, c(2:9, 20, 28:31, 36:45)]
write.csv(ratio_transformed_zhenzhen,
          "/Users/zhenzhenzhu/Documents/capstone/capstone_repository/data/ratio_transformed_zhenzhen.csv")
```



## Correlation
```{r}
colnames(tech)
cor(subset(tech[,9:32][sapply(tech[,9:32], is.numeric)], select = -Return3), tech$Return3)
##log
cor(log(tech$Earning.Per.Share), tech$Return, use = 'complete.obs') #

```

## linear regression with original variables
```{r}
tech_num <- tech[,9:32]
m1 <- lm(Return ~., data=tech_num)
summary(m1)
```

## linear reg with log transformed Return on Assets 
```{r}
colnames(tech)
```



